Feature Selection and Classification of \\ Spam on Social Networking Sites

### 2012-05-09 docs/paper/paper.tex
update

\documentclass[preprint]{acm_proc_article-sp}
%\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{graphicx,subfigure}
\usepackage{xspace}
\usepackage{multirow}

\newcommand{\ie}{{\em i.e.,}~}
\newcommand{\eg}{{\em e.g.,}~}

@@

Social networks of any significant size see a constant stream of spam, scams and phishing attacks. 
The nature of this unwanted activity can be quite diverse, specific to a site's target audience and 
often not trivial to detect. Marketers can spam members with unwanted advertisements, fraudsters 
lure users with advance fee frauds and other confidence tricks, while others may attempt to steal user
information by directing users to external phishing pages. In contrast to email spam, social spam 
has a more personal component, since both spammers and legitimate users (``ham" users) have accompanying 
profiles with descriptive information.

@@

We investigated several machine learning techniques to automatically 
detect spam and scams in the private message data, and chose to use 
train and evaluate Na\"ive Bayes, Linear Regression, and 
Support Vector Machine (SVM) classifiers. Our implementations used 
a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
an in-memory distributed computing framework designed for machine 
learning and iterative computation.

@@

The rapid growth of social media has made Social Networking Services (SNSs) 
increasingly attractive targets for spam and fraud, leading to a proliferation 
of sophisticated attacks. This trend is reflected in recent research, as papers 
have focused on identifying and classifying the various types of social media spam. 
Many of these studies employ techniques previously used to combat conventional 
email and web spam. SNSs also provide opportunities to take advantage of user 
reputation and other social graph-dependent features to improve classification. 
Nevertheless, most research has been carried out on publicly-available data 
from SNSs, making it difficult up until now to measure the effect of private user 
data on algorithms for detecting site misuse.

@@

We did not use all of the expert features available to us, due to time constraints. Likewise, for many expert features could 
have a wide range of values, which would expand the dictionary size considerably. In addition to the bag-of-words from the 
message text, we used the following features:
\begin{itemize}
\item Sex of sender
\item Age of sender
\item Age of recipient
\item Account lifetime
\item Month of birth 
\item Whether sender has friends , boolean
\item Whether sender has photo(s) , boolean
\item Profile country and IP-detected country match , boolean
\item IP-detected country is Ghana, Nigeria, Senegal, Malaysia, Turkey, Gambia, Ivory Coast, Togo (``high-risk" countries) , boolean
\end{itemize}

@@

% Modified from Reynold's HW1  write-up
%We use a na\``ive Bayes model, in which the documents are generated according to the following process.  
%First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
%Then, a document from class \(c\) is generated.
%We tried a multinomial model and a Bernoulli model to model this process.
%\begin{itemize}
%\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
%\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
%\end{itemize}
%The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.
%
%

@@

One of the advantages of the Na\"ive Bayes classifier was that it took very little time to train the classifier, 
and the steps involved were all trivially parallelizable, which makes it convenient for large data sets.

The decline in accuracy after adding expert features is most likely explained by the ``naive" assumption in this 
algorithm that the features are independent. Several of the features that we used, \eg{account lifetime, 
and whether a user has friends or photos}, are in fact likely to be co-dependent, leading to an overweight of 
these features during classification. The wealth of expert features available to a SNS, many of which are 
not independent, makes this the least compelling classifier choice that we examined.

@@

Our initial results using the Pegasos SVM solver included in ScalaNLP were lower than expected. The 
classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
51\%. Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
feature models. We see that for the bag-of-words classifier, the words ``year," ``new," ``happy,"  
``2012," ``wish," and ``you" all have high weights, which makes sense given that the ham data was 
collected around the beginning of 2012. The table shows that three expert features: account lifetime, 
friends and recipient age, are weighted very heavily in comparison to the remaining features. These 
skewed weights (and the inability to tune the SVM solver via ScalaNLP) prompted us to test LIBSVM 
on the same data.

@@

year & 0.2253 & F-account\_lifetime & -0.9800\\
new & 0.2250 & F-friends & -0.3671\\
am & -0.1722 & F-recip\_age & -0.2129\\
happy & 0.1600 & F-age & -0.1830\\
d & 0.1513 & i & -0.05231\\
ok & -0.1497 & you & -0.0483\\
yahoo & -0.1360 & to & -0.04026\\
it & 0.1200 & and & -0.03636\\
me & -0.1180 & F-month\_of\_birth & -0.03452\\
too & 0.1173 & F-country\_match & -0.0269\\
i'm & 0.1116 & me & -0.0224\\
haha & 0.0998 & am & -0.0220\\
can & -0.0971 & my & -0.0209\\
okay & -0.0947 & ? & 0.0199\\
dear & -0.0882 & ? & 0.0191\\
p & 0.0862 & a & -0.01573\\
here & -0.0844 & your & -0.01554\\
2012 & 0.084 & F-photo\_exists & -0.0146\\
wish & 0.0836 & the & -0.01412\\
doing & -0.0779 & f-sex & -0.01258\\

@@

Our results with LIBSVM's RBF kernel were consistently better than any of the other classifiers we tested. We 
evaluated the classifier's performance on five pairs of bag-of-words and expert feature data sets ranging in 
size from 2000 to 100k messages. Table \ref{tab:svmacc} shows results of 10-fold cross-validation. Our mean 
accuracy over the bag-of-words data sets was 85.73\% and 98.41\% over the bag-of-words with expert features. We 
observed a slight decrease in performance on the bag-of-words data sets as the data size grew. Figure \ref{fig:roc-25}
shows the AUCs and ROC curves for the 50k data sets, making clear the the significantly greater AUC resulting 
from the addition of expert features.

@@

In this paper, we presented a brief overview of the breakdown of spam and scams on the InterPals web site. 
We then examined a number of statistics on both public and private data that showed that spam users, even when 
grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
them from legitimate users. To simulate an approach 
These features allowed us to improve the accuracy of classifiers

%\begin{figure}[h]
%    \centering
%    \includegraphics[width=\linewidth]{figures/sex.pdf}
%    \caption{Sender sex as stated on profile.}
%    \label{fig:sendsex}
%\end{figure}
%

>>>>

\documentclass[preprint]{acm_proc_article-sp}
%\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{graphicx,subfigure}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{amsmath}

\DeclareMathOperator{\Argmax}{argmax}
\newcommand{\PosC}{\mathrm{Spam}}
\newcommand{\NegC}{\mathrm{Ham}}
\newcommand{\ie}{{\em i.e.,}~}
\newcommand{\eg}{{\em e.g.,}~}

@@

Social networks of any significant size witness a constant flow of spam, scams and phishing attacks. 
The nature of this unwanted activity can be quite diverse, specific to a site's target audience and 
often not trivial to detect. Marketers can spam members with unwanted advertisements, fraudsters 
lure users with advance fee frauds and other confidence tricks, while others may attempt to steal user
information by directing users to external phishing pages. In contrast to email spam, social spam 
has a more personal component, since both spammers and legitimate users (``ham" users) have accompanying 
profiles with descriptive information.

@@

We investigate several machine learning techniques to automatically 
detect spam and scams in the private message data, and train and 
evaluate Na\"ive Bayes, Linear Regression, and Support Vector Machine (SVM) 
classifiers. Our implementations used a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
an in-memory distributed computing framework designed for machine learning and iterative computation.

@@

The rapid growth of social media has made SNSs increasingly attractive targets 
for spam and fraud, leading to a proliferation of sophisticated attacks. This 
trend is reflected in recent research, as papers 
have focused on identifying and classifying the various types of social media spam. 
Many of these studies employ techniques previously used to combat conventional 
email and web spam. SNSs also provide opportunities to take advantage of user 
reputation and other social graph-dependent features to improve classification. 
Nevertheless, most research has been carried out on publicly-available data 
from SNSs, making it difficult up until now to measure the effect of private user 
data on algorithms for detecting site misuse.

@@

We did not use all of the expert features available to us, due to time constraints. Likewise, for many expert features could 
have a wide range of values, which would expand the dictionary size considerably. In addition to the bag-of-words from the 
message text, we used the following features:
\begin{itemize}
\item Sex of sender
\item Age of sender
\item Age of recipient
\item Account lifetime
\item Month of birth 
\item Whether sender has friends, boolean
\item Whether sender has photo(s), boolean
\item Profile country and IP-detected country match, boolean
\item IP-detected country is a ``high-risk"\footnote{As determined by us for the purposes of this project, based on 
the IP-detected country histogram of spam users. This list includes Ghana, Nigeria, Senegal, Malaysia, Turkey, 
Gambia, Ivory Coast and Togo.} spam country, boolean
\end{itemize}

@@

The main idea of the Naive Bayes classification algorithm is that one can calculate the probability 
of an item belonging to a certain class by applying Bayes' theorem on the assumption that the probability 
of one word/feature that appears in a document is independent of the probability of any other feature. 
While is very rarely the case (hence ``naive"), Naive Bayes still performs surprisingly well in many cases.

Using the naive Bayes classifier, for a class \(c \in \{\PosC, \NegC\}\) and document \(d\) consisting of words \(w_{d,1},\dotsc,w_{d,m_d}\) (where each word occurs at most once in the Bernoulli version), we have
\[\Pr(c|d)=\frac{\Pr(d|c)\Pr(c)}{\Pr(d)}\]
where
\[\Pr(d|c)=\sum_{i=1}^{m_d}\Pr(w_i|c)\ldotp\]
The training procedure sets \(\Pr(c)\) to be the fraction of documents with class \(c\),%

\[\Pr(w|c)=\frac{n_{w,c}+\alpha}{\sum_{w'}(n_{w',c}+\alpha)}\ldotp\]
When testing, our model classifies a document \(d\) as \(\Argmax_c\Pr(c|d)\).

%We use a na\``ive Bayes model, in which the documents are generated according to the following process.  
%First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
%Then, a document from class \(c\) is generated.
%We tried a multinomial model and a Bernoulli model to model this process.
%\begin{itemize}
%\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
%\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
%\end{itemize}
%The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.
%
%

@@

The Na\"ive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
the features that we used, \eg{account lifetime, and whether a user has friends or photos}, are in 
fact likely to be co-dependent, leading to an overweight of these features during classification.

One of the advantages of the Na\"ive Bayes classifier was that it took very little time to train the classifier, 
and the steps involved were all trivially parallelizable, which makes it convenient for large data sets. 
However, in view of the wealth of expert features available to a SNS, many of which are not independent, this was 
the least compelling classifier choice that we examined.

@@

Our initial results using the Pegasos SVM solver included in ScalaNLP were lower than expected. The 
classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
51\%.

Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
feature models. We see that for the bag-of-words classifier, the words ``year," ``new," ``happy,"  
``2012," ``wish," and ``you" all have high weights, which makes sense given that the ham data was 
collected around the beginning of 2012. The lists include a number of stopwords, such as ``ok," 
``me" and ``it," which should ideally not be highly weighted. While we did not remove stopwords for this 
project, this is something that we could consider doing in the future.

The table of top terms reveals that three expert features: account lifetime, friends and 
recipient age, are weighted very heavily in comparison to the remaining features. These 
skewed weights (and the inability to tune the SVM solver via ScalaNLP) prompted us to test LIBSVM 
on the same data.

@@

year & 0.2253 & F-account\_lifetime & -0.9800\\
new & 0.2250 & F-friends & -0.3671\\
am & -0.1722 & F-recip\_age & -0.2129\\
happy & 0.1600 & F-age & -0.1830\\
d & 0.1513 & i & -0.05231\\
ok & -0.1497 & you & -0.0483\\
yahoo & -0.1360 & to & -0.04026\\
it & 0.1200 & and & -0.03636\\
me & -0.1180 & F-month\_of\_birth & -0.03452\\
too & 0.1173 & F-country\_match & -0.0269\\
i'm & 0.1116 & me & -0.0224\\
haha & 0.0998 & am & -0.0220\\
can & -0.0971 & my & -0.0209\\
okay & -0.0947 & ? & 0.0199\\
dear & -0.0882 & ? & 0.0191\\
p & 0.0862 & a & -0.01573\\
here & -0.0844 & your & -0.01554\\
2012 & 0.084 & F-photo\_exists & -0.0146\\
wish & 0.0836 & the & -0.01412\\
doing & -0.0779 & F-sex & -0.01258\\

@@

Our results with LIBSVM's RBF kernel were consistently better than any of the other classifiers we tested. We 
evaluated the classifier's performance on five pairs of bag-of-words and expert feature data sets ranging in 
size from 2000 to 100k messages. Table \ref{tab:svmacc} shows results of 10-fold cross-validation. Our mean 
accuracy over the bag-of-words data sets was 85.73\% and 98.41\% over the bag-of-words with expert features. We 
observed a slight decrease in performance on the bag-of-words data sets as the data size grew. Figure \ref{fig:roc-25}
shows the AUCs and ROC curves for the 50k data sets, making clear the the significantly greater AUC resulting 
from the addition of expert features. As expected, however, both the ScalaNLP and LIBSVM implementations 
took substantially longer to train than NB and LR.

@@

In this paper, we presented a brief overview of the breakdown of spam and scams on the InterPals web site. 
We then examined a number of statistics on both public and private data that showed that spam users, even when 
grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
them from legitimate users. We trained and evaluated Na\"ive Bayes, Linear Regression and SVM classifiers on a subset 
of the 4 million messages on which we collected feature statistics. Our evaluation compared the performance of 
each classifier on bag-of-word representations of message text vis-\`{a}-vis the combination of bag-of-word features 
with a number of ``expert features." We found that SVMs using an RBF kernel on LIBSVM outperformed the other 
classifiers. In both the linear regression and SVM cases, the addition of expert features substantially increased 
the classifier's accuracy, underscoring the impact of supplementary account-related features in detecting spam on 
SNSs.
